---
title: "HydroBasin: End‑to‑End Workflow"
author: "Anugraha Das"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HydroBasin: End-to-End Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  message  = FALSE,
  warning  = FALSE
)
```  

## Introduction  
`HydroBasin` streamlines two labor‑intensive hydrology tasks:

1. **Watershed delineation** from an area‑of‑interest polygon.
2. **Monthly water‑balance/run‑off** calculation using TerraClim climate grids.

This vignette shows an end‑to‑end example you can run on any machine with an
internet connection.  The *heavy* steps are wrapped in `eval = FALSE` code
chunks so knitting never stalls when CRAN or CI checks are offline.

## Load the package  
```{r load}
library(HydroBasin)
```

If you installed HydroBasin from GitHub, be sure the attached version is the
one you just built:
```{r, eval = FALSE}
packageVersion("HydroBasin")
```

---

## 1  Delineate a basin  
We’ll start with the demo AOI shapefile bundled in the package.

```{r basin, eval = FALSE}
# Path to example extent
aoi <- system.file("extdata/aoi_example.shp", package = "HydroBasin")

# Delineate (interactive pour‑point picker will pop up)
basin <- delineate_basin(
  aoi      = aoi,
  out_dir  = file.path(tempdir(), "hb_demo"),
  quiet    = TRUE
)

# Inspect
basin
```

> **Tip** Once you are happy with an outlet location, save the generated
> `pour_point.shp` and feed it back to `delineate_basin()` next time to run
> fully non‑interactive.

---

## 2  Download climate grids & compute run‑off  
Using the freshly delineated boundary we pull TerraClim
precipitation (**ppt**) and actual evapotranspiration (**aet**) for
2020–2024, convert millimetres to cubic metres, and calculate run‑off.

```{r runoff, eval = FALSE}
runoff <- calculate_runoff(
  basin       = basin,           # sf object from step 1
  start_date  = "2020-01-01",
  end_date    = "2024-03-31",
  return_plot = TRUE,
  quiet       = TRUE
)

# First six rows
head(runoff$data)

# Plot (returns a ggplot object)
runoff$plot
```

<small>*Note*  The chunk is `eval = FALSE` to keep vignette build time low.
To run locally, set `eval = TRUE` or paste the code in your R console.</small>

---

## 3  Persist the results  
```{r write, eval = FALSE}
readr::write_csv(runoff$data, "monthly_basin_runoff_volume.csv")
```

The CSV has one row per month with total precipitation, AET and run‑off
volumes (m³) plus convenient `year` and `month` columns for aggregation.

---

## 4  Automation tips  
* **Batch processing** Loop over multiple AOIs or pour‑points by storing them
  in a data frame and calling the two functions inside `purrr::map()`.
* **Fixed outlet** Use the `pour_pt` argument in `delineate_basin()` to
  supply a stored pour‑point and skip the interactive step.
* **Coarser DEM** For huge regions set `dem_zoom = 10` to download ~90 m tiles
  and speed up WhiteboxTools.

---

## Session info  
```{r session}
sessionInfo()
```

